### 2014-06-06 paper/body.tex
minimum necessary figure caption
writing edits

(((( two gabor features from very different networks ))))

@@

Many trained networks reported recently in the literature --- unsupervised and supervised alike --- exhibit a curious aspect in common: they all learn features that resemble gabor filters on the first layer (see Figure~\ref{fig:layer1}) \editbox{fix broken figure ref here}. This happens with such frequency that not only is it no longer surprising, but to obtain anything else on a natural-image dataset causes suspicion of poorly chosen hyperparameters or a bug somewhere in one's code. In fact, this paper's authors are now used to checking for gabor features in the first layer simply to assure ourselves that the training is at least partially working.

Because finding gabor features on the first layer seems to occur regardless of the exact cost function and natural image dataset used, we may call these these features \emph{general}. On the other hand, we know that the last layer of a trained network will depend greatly on the chosen \{network model, dataset\} combination. For example, for networks trained toward a supervised classification objective, each output unit will be specific to a particular class, at least to the extent that training is successful. In this sense, let's say that the last layer features are \emph{specific}. So far these terms are only intuitively defined, but we shall give one possible more rigorous definition soon.

If the layer 1 features are general, and the last layer features are specific, we know that somewhere in the network there must be a transition from general to specific. This observation raises a few questions:

@@

We have noted the curious tendency of gabor features to show up in the first layer of neural networks trained on natural images for a large range of network architectures and cost functions. We say these features are general if they are useful for many different tasks or specific if they are useful only for a few. We can measure the specificity of a set of features by training it on one task \dA and then using it for another task \dB.

@@

\begin{figure}[htpb]
\begin{center}
\includegraphics[width=1\linewidth]{plots/result_transfer_crop.pdf}
\includegraphics[width=1\linewidth]{plots/result_transfer_lines_crop.pdf}
\end{center}
\caption{\edit{Caption here...}}
\label{fig:results}
\end{figure}

@@

\item The dark blue \net{AnA} points show a curious behavior. We can see at layer one that the performance is the same (as the \net{baseA} points), which was expected. If we learn eight layers of features, save the first layer of learned gabor features, and reinitialize the whole network and retrain it toward the same task, we'd expect to do just as well. This is also true for layer 2. Layers 3, 4, 5, and 6, particularly 4 and 5, behave somewhat differently. When we keep the first, say, 5 layers and retrain layers 6-8, performance suffers. This is evidence that the original network contained \emph{co-adapted features} on successive layers, that is features that interact with each other in a complex way such that this co-adaptation \emph{could not be relearned} by the upper layers. In other words, gradient descent was able to find a good solution the first time, but this was only possible because the layers were jointly trained. As we continue to the right, we see that layer 6 is nearly back to the base level of performance, as is layer 7. As we get closer and closer to the final 500-way softmax output layer (8), there is less to relearn, and apparently relearning these one or two layers is simple enough for gradient descent to find a good solution. Alternately, we may say that there is less co-adaptation of features between layers 6 and 7 or between 7 and 8.

@@

\item The dark red \net{BnA} points show the effect we set out to measure in the first place: the transferability of features from one network to another at each layer. Layers one and two transfer almost perfectly from \net{B} to  \net{A}, giving evidence that at least for our two tasks, not only are the gabor features general, but the second layer features are as well. Layer three shows a slight drop, and layers 4-7 show a more significant drop in performance. Thanks to the \net{ANA} points, we can tell that this drop is from a combination of two separate effects: the drop from lost co-adaptation \emph{and} the drop from features which are less and less general. On layers 3, 4, and 5, the first effect dominates, whereas on layers 6 and 7 the first effect diminishes and the specificity of representation dominates the drop in performance.

>>>>

(((( two Gabor features from very different networks ))))

@@

Many trained networks reported recently in the literature --- unsupervised and supervised alike --- exhibit a curious aspect in common: they all learn features that resemble Gabor filters in the first layer (see Figure~\ref{fig:layer1}) \todo{fix broken figure ref here}. The appearance of these Gabor filters is so common that to obtain anything else on a natural-image dataset causes suspicion of poorly chosen hyperparameters or a software bug. 
%In fact, this paper's authors are now used to checking for Gabor features in the first layer simply to assure ourselves that the training is at least partially working.

Because finding Gabor features on the first layer seems to occur regardless of the exact cost function and natural image dataset used, we may call these these features \emph{general}. On the other hand, we know that the last layer of a trained network will depend greatly on the chosen \{network model, dataset\} combination. For example, for networks trained toward a supervised classification objective, each output unit will be specific to a particular class, at least to the extent that training is successful. In this sense, let's say that the last layer features are \emph{specific}. So far these terms are only intuitively defined, but we shall provide a candidate for a more rigorous definition below. \todo{Consider something like: ``Empirically, we know that these filter are generic because they can be transferred to other networks and perform well [cites]. Transferring information learned on one task or dataset to a new task or data set is know as \emph{Transfer Learning} [cites].''}

If these Gabor features of the first features are general, and the features of the last layer are specific, there must be a transition from general to specific somewhere in the network. This observation raises a few questions:

@@

We have noted the curious tendency of Gabor features to show up in the first layer of neural networks trained on natural images for a large range of network architectures and cost functions. We say these features are general if they are useful for many different tasks or specific if they are useful only for a few. We can measure the specificity of a set of features by training it on one task \dA and then using it for another task \dB.

@@

\begin{figure}[htpb]
\begin{center}
\includegraphics[width=1\linewidth]{plots/result_transfer_crop.pdf}
\includegraphics[width=1\linewidth]{plots/result_transfer_lines_crop.pdf}
\end{center}
\caption{Accuracy of each treatment and control. \emph{Top panel:} Raw data. \emph{Bottom panel}: Lines connect the medians of each treatment. Shaded areas and text highlight key takeaway messages. \todo{This is the min. required if we explain the figure in the main text.}
}
\label{fig:results}
\end{figure}

@@

\item The dark blue \net{AnA} points show a curious behavior. We can see at layer one that the performance is the same (as the \net{baseA} points), which was expected. If we learn eight layers of features, save the first layer of learned Gabor features, and reinitialize the whole network and retrain it toward the same task, we'd expect to do just as well. This is also true for layer 2. Layers 3, 4, 5, and 6, particularly 4 and 5, behave somewhat differently. When we keep the first, say, 5 layers and retrain layers 6-8, performance suffers. This is evidence that the original network contained \emph{co-adapted features} on successive layers, that is features that interact with each other in a complex way such that this co-adaptation \emph{could not be relearned} by the upper layers. In other words, gradient descent was able to find a good solution the first time, but this was only possible because the layers were jointly trained. As we continue to the right, we see that layer 6 is nearly back to the base level of performance, as is layer 7. As we get closer and closer to the final 500-way softmax output layer (8), there is less to relearn, and apparently relearning these one or two layers is simple enough for gradient descent to find a good solution. Alternately, we may say that there is less co-adaptation of features between layers 6 and 7 or between 7 and 8.

@@

\item The dark red \net{BnA} points show the effect we set out to measure in the first place: the transferability of features from one network to another at each layer. Layers one and two transfer almost perfectly from \net{B} to  \net{A}, giving evidence that at least for our two tasks, not only are the Gabor features general, but the second layer features are as well. Layer three shows a slight drop, and layers 4-7 show a more significant drop in performance. Thanks to the \net{ANA} points, we can tell that this drop is from a combination of two separate effects: the drop from lost co-adaptation \emph{and} the drop from features which are less and less general. On layers 3, 4, and 5, the first effect dominates, whereas on layers 6 and 7 the first effect diminishes and the specificity of representation dominates the drop in performance.
